### 2015-06-13 ApresentacoesArtigos/quali/quali.tex
trabalho nos metodos

\begin{center}
\Huge{Complex networks for the participant}
\end{center}

@@

Área de concentração: Física Básica

@@

\begin{itemize}
	\item Redes do Facebook: 8 são usadas como referência em~\cite{timeS}, mas dezenas, talvez algumas centenas, foram observadas nos experimentos da Seção~\ref{sec:exp}.
	\item milhares de tweets (talvez alguns milhões), dados de portais
 	\item mecanismos participativos como o Participa.br, Cidade Democrática e o AA.
\end{itemize}

\section{Minhas próprias redes sociais}
Considerações sobre o uso de minhas anotações para
pesquisa sobre mim e meu mundo.

\chapter{Métodos}
\section{Estatística circular}
Messages were counted along time with respect to seconds, minutes, hours, days of the week, days of the month, and months of the year. This resulted in histograms from which patterns could be drawn. The ratio $\frac{b_h}{b_l}$ between the highest and lowest incidences on the histograms served as a hint of how close the observed distribution is to a uniform distribution.

The average and the dispersion were taken using circular statistics, in which each $measurement$ (data point) is represented as a complex number with modulus equal to one, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, where $\theta=measurement\frac{2\pi}{period}$. The moments $m_n$, lengths of moments $R_n$, mean angle $\theta_\mu$, and rescaled mean angle $\theta_\mu'$ are defined as:

\begin{align}\label{eq:cmom}
	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
	R_n&=|m_n|\\
	\theta_\mu&=Arg(m_1) \nonumber \\
	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
\end{align}

$\theta_\mu'$ is used as the measure of location. Dispersion is measured using the circular variance $Var(z)$, the circular standard deviation $S(z)$, and the circular dispersion $\delta(z)$:

@@

As expected, a positive correlation was found in all $Var(z)$, $S(z)$ and $\delta(z)$ dispersion measures, as can be noticed in Section~\ref{si:circ} of the Supporting Information, and $\delta(z)$ was preferred in the discussion of results.

\subsection{Formação das redes de interação}\label{sec:intNet}
Interaction networks can be modeled both weighted or unweighted, both directed or undirected~\cite{bird,newmanCommunityDirected,newmanCommunity2013}.
Networks in this article are directed and weighted, the more informative of trivial possibilities, i.e. we did not investigate directed unweighted, undirected weighted, and undirected unweighted representations of the interaction networks. 
The networks were obtained as follows: a direct response from participant B to a message from participant A yields an edge from A to B, as information went from A to B. The reasoning is: if B wrote a response to a message from A, he/she read what A wrote and formulated a response, so B assimilated information from A, thus $A \rightarrow B$. Inverting edge direction yields the status network: B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$. This article uses the information network as described above and depicted in Figure~\ref{formationNetwork}. Edges in both directions are allowed. Each time an interaction occurs, one is added to the edge weight. Selfloops were regarded as non-informative and discarded. These human social interaction networks are reported in the literature as exhibiting scale-free and small world properties, as expected for (some) social networks~\cite{bird,newmanBook}.

@@

@@

%\chapter{Finished and planed tasks, chronogram}
\chapter{Afazeres finalizados, planejados e prontos; cronograma}
% Colocar tabela com os comparativos
% Referenciar para complementos no texto.
\section{Experimentos}\label{sec:exp}
\section{Documents}
\subsection{To be finished}
\subsubsection{Anthropological physics}
The study of human systems raises conceptual
and ethical issues that require anthropological considerations.
There are two immediate routes to this concepts:
\begin{itemize}
	\item What data should or can be used?
	\item Can one experiment in a network of humans? In which context?
\end{itemize}

>>>>

\begin{center}
%\Huge{Complex networks for the participant}
\Huge{Redes complexas para o participante}
\end{center}

@@

Área de concentração: Física Aplicada
Opção: Física Computacional

@@

\begin{itemize}
	\item Redes de amizade e interação do Facebook. 8 são usadas como referência em~\cite{timeS}, mas dezenas, talvez algumas centenas, foram observadas nos experimentos da Seção~\ref{sec:exp}.
	\item Milhares de tweets (talvez alguns milhões), geralmente vinculados à alguma \emph{hashtag}. Em especial, a rede de retweets de 22 mil tweets com a hashtag \#arenaNETmundial, foi analizada em~\cite{timeS}.
	\item Mecanismos participativos como o Participa.br, Cidade Democrática e o AA. As redes de amizade e de interação do Participa.br foram analizadas em~\cite{timeS}.
\end{itemize}

\chapter{Métodos}
\section{Estatística temporal e circular}

Para observação de padrões temporais, foram consideradas escalas diferentes.
Em cada escala, de segundos e meses, foram construídos histogramas de atividade
e feitas algumas medidas de estatística circular.
A fração $\frac{b_h}{b_l}$ entre a maior $b_h $ e a menor $b_l$ incidência nos histogramas
serviram como pista sobre quão uniforme são as distribuições observadas.

Considere cada $ medida $ (dado pontual) como um número complexo com módulo 1, $z=e^{i\theta}=\cos(\theta)+i\sin(\theta)$, onde $\theta= medida \frac{2\pi}{periodo}$. Os momentos $m_n$, tamanhos dos momentos $R_n$, ângulo médio $\theta_\mu$, e o ângulo médio reescalado $\theta_\mu'$ são definidos assim:
\begin{align}\label{eq:cmom}
	m_n&=\frac{1}{N}\sum_{i=1}^N z_i^n \nonumber\\
	R_n&=|m_n|\\
	\theta_\mu&=Arg(m_1) \nonumber \\
	\theta_\mu'&=\frac{period}{2\pi} \theta_\mu \nonumber
\end{align}

$\theta_\mu'$ é usado como medida de localização. A dispersão é medida usando a variância circular $Var(z)$, 
o desvio padrão circular $S(z)$, e a dispersão circular $\delta(z)$:

@@

Como esperado, há uma correlação positiva entre $Var(z)$, $S(z)$ e $\delta(z)$,
como pode ser notado nas informações de suporte de~\cite{timeS}.
A medida $\delta(z)$ foi preferida na discussão dos resultados.

\subsection{Formação das redes de interação}\label{sec:intNet}
Redes de interação podem ser modeladas tanto com quanto sem peso, tanto dirigida quando não dirigida~\cite{bird,newmanCommunityDirected,newmanCommunity2013,newmanBook}.
Neste trabalho, quando possível, consideramos redes dirigidas e com peso, a mais informativa das possibilidades.
Nestes casos, desconsideramos as versões dirigidas sem peso, não dirigidas com peso e não dirigidas e sem peso.

Em geral, as redes de interação são obtidas da seguinte forma:
uma reação direta do participante B a uma mensagem do participante A implica em uma aresta de A para B,
representando a informação que foi de A para B.
O raciocínio é: se B reagiu
a uma mensagem de A, ele/ela leu o que A escreveu e formulou uma reação, portanto B
assimilou informação de A, assim $A \rightarrow B$.
A inversão da direção da aresta produz a rede de status: B leu a mensagem e considerou
o que A escreveu digno de resposta, dando status para A, portanto $B \rightarrow A$.
Neste trabalho, as redes de interação são dirigida conforme o fluxo de informação, $A \rightarrow B$.
A Figura~\ref{formationNetwork} expõe esta formação. Maiores detalhes são:

Inverting edge direction yields the status network: B read the message and considered what A wrote worth responding, giving status to A, thus $B\rightarrow A$. This article uses the information network as described above and depicted in Figure~\ref{formationNetwork}. Edges in both directions are allowed. Each time an interaction occurs, one is added to the edge weight. Selfloops were regarded as non-informative and discarded. These human social interaction networks are reported in the literature as exhibiting scale-free and small world properties, as expected for (some) social networks~\cite{bird,newmanBook}..

@@

% explicitar que a regra de formação não é só para mensagens de email

@@

%\chapter{Finished and planed tasks, chronogram}
\chapter{Afazeres e cronograma}
% Colocar tabela com os comparativos
% Referenciar para complementos no texto.
\section{Experimentos}\label{sec:exp}
\section{Documents}
\subsection{To be finished}
\subsubsection{Anthropological physics}
The study of human systems raises conceptual
and ethical issues that require anthropological considerations.
There are two immediate routes to this concepts:
\begin{itemize}
	\item What data should or can be used?
	\item Can one experiment in a network of humans? In which context?
\end{itemize}
